---
title: "Graphs in R"
output: html_notebook
---

# We're going to cover:

 - ggplot()
    - scatter plolts
    - line plots
    - boxplots
    - details
        - titles
        - colors
        - facettes
        - multi-figures


We're going to use the library wooldridge and import the ceo salary dataset. This is a sneakpeak on the datasets that we will be using in eco 585: applied econometrics.

It create a dataframe called ceosal1

```{r}
library(wooldridge)

data("ceosal1")

names(ceosal1)
```

We're going to do some simple examples of graphing with the premiere graphing package in R - ggplot2. This is included in tidyverse, so let's load that in

```{r}
library(tidyverse)
```

## Scatterplot

Your traditional scatter plots. You must tell R the data set you are pulling, the x and y axis and then add , that's right, add (plus sign) that you would like a scatter plot. This is particularly important for continuous variables.

```{r}
ggplot(data = ceosal1, mapping = aes(x = lsalary, y =  lsales)) + geom_point()
```

What's kind of nice about this is that you can assign the core information to a variable and then add on what kind of graph you want to see.
```{r}
sal_sales <- ggplot(data = ceosal1, mapping = aes(x = lsalary, y =  lsales)) 

#below is where you see that additional graphing capability
sal_sales + geom_point()
```

# Lineplots


So, creating line graphs are easy. We just change the ending of the type of plot we want. 
```{r}
sal_sales + geom_line()
```

## boxplots
box plots are similarly easy to show

```{r}

ggplot(data = ceosal1, mapping = aes(y = lsalary))  + geom_boxplot()

 # we can even add some more detail about the distribution and change the color
sal_sales + geom_boxplot() + geom_jitter(alpha = 0.3, color = "tomato")

```
```{r}

#creating a category variable that I can get information by groups
ceosal <- pivot_longer(ceosal1, c('indus':'utility'), names_to='cat_industry', values_to= 'roe2')
ceosal<-subset(ceosal,  roe2!=0)
ceosal$roe2<- NULL

```


```{r}
ggplot(data = ceosal, mapping = aes(y = lsalary, x  = cat_industry)) + geom_boxplot() 
```

## Faceting

You can create multiple graphs in one figure through faceting. 

```{r}
ceo_graph <- ggplot(data = ceosal, mapping = aes(y = lsalary, x  = lsales)) 

ceo_graph + geom_line()  + facet_wrap(facets = vars(cat_industry))
```
# Customization

We can do a lot of custom details to make our graphs look perfect.  You can take a look at different themes[here](https://ggplot2.tidyverse.org/reference/ggtheme.html).

```{r}
ceo_sal_line <- ceo_graph + geom_line() +
labs(title = "Relationship of Ceo Salary and Sales by Industry",
        x = "Salary (logged)",
        y = "Sales (logged)") +
    theme_bw() +  # we can add themes (same as in python)
    theme(text=element_text(size = 16))  #specify text size across the entire graph

ceo_sal_line
```

Some good advice on how to write short descriptions about data visualization can be found here: 
https://medium.com/nightingale/writing-alt-text-for-data-visualization-2a218ef43f81 

This is really for thinking about alt text, what we often do for twitter or even markdown - but I think it's a good way to think about how to title and explain graphs quickly.

## Colors

As always we can add colors. Color is an option in any of the graphs funciton we use. Lots of pallettes, colors and options can be found [here](https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/).    

```{r}
library(RColorBrewer) # This package helps add color palettes

ggplot(data = ceosal, mapping = aes(y = lsalary, x  = cat_industry)) + geom_point(aes(color = cat_industry))

```

We can also add our transparency 

```{r}
sal_sales + geom_point(alpha = .3)
```



# Multiple Plots
As in python, you can put together plots on one figure

```{r}
#must import this package to facilitate (don't forget to install it - only have to do it once)
library(patchwork)

boxplots <- ggplot(data = ceosal, mapping = aes(y = lsalary, x  = cat_industry)) + geom_boxplot() +
  labs(y = "Salary (logged)",
        x = "Categories")

ceo_sal_line / boxplots + plot_layout(heights = c(2, 2))
```

# exporting
with the funciton ggsave, you can export any graph (remember it's where you set your working directory unless you specify otherwise)
```{r}

plots <- ceo_sal_line / boxplots + plot_layout(heights = c(2, 2))

ggsave("2plots.png", plots, dpi = 300)
#saving at a dpi of 300
# you can also specify the height and width if needed
```

